---
title: "Allegations Against the NYPD"
author: "James P. Hare"
date: "7/28/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(sf)

#Police precinct map from NYC Open Data https://data.cityofnewyork.us/Public-Safety/Police-Precincts/78dh-3ptz#revert
nyc_precincts <- read_sf("data/Police Precincts/geo_export_cfbe857e-58ec-4848-b99b-1f33fd967621.shp")

#Alleged misconduct from ProPublica
allegations <- read_csv("data/allegations_20200726939.csv")

```

Following a change to the New York law that kept police officers' disciplinary records secret---and amid an ongoing lawsuit---ProPublica has released a [searchable database](https://projects.propublica.org/nypd-ccrb/) of complaints to to the Civilian Complaint Review Board (CCRB). The database, which can be downloaded in its entirety [here](https://www.propublica.org/datastore/dataset/civilian-complaints-against-new-york-city-police-officers), includes allegations against the nearly 4,000 officers who have at least one substantiated complaint against them. The CCRB's powers are extremely circumscribed, and these data reflect its [limitations](https://www.propublica.org/article/nypd-civilian-complaint-review-board-editors-note). The CCRB "exonerates" officers whose conduct is ruled to fall within departmental guidelines, no matter how egregious. Allegations may remain "unsubstantiated" due to a routine lack of NYPD cooperation (in violation of the law), and even "substantiated" allegations lead only to suggestions, which the department is free to ignore. 

Even a quick glance through these records turns up numerous officers who have faced repeated, serious, substantiated misconduct allegations while rising through the ranks. But looking at these data from a bird's eye view also reveals some striking patterns in the complaints against NYPD misconduct. Anyone who is familiar with New York City would not expect police misconduct---and thus complaints about police misconduct---to be evenly distributed geographically, and that is the case here. 

```{r}
precinct_allegations <- allegations %>%
        filter(precinct < 124 & precinct > 0 & 
                       board_disposition != "Unsubstantiated" &
                       board_disposition!= "Exonerated") %>%
        group_by(precinct) %>%
        summarise(n = n()) %>%
        left_join(nyc_precincts)

ggplot(precinct_allegations) +
        geom_sf(aes(fill = n, geometry = geometry)) +
        labs(title = "Substantiated Complaints Against NYPD by Precinct",
             subtitle = "1985 to 2020",
             caption = "Source: ProPublica") +
        theme(legend.position = c(0, 1),
              legend.justification = c(-.4, 1.1),
              legend.title = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(),
              panel.background = element_blank()) +
        scale_fill_continuous(type = "viridis")
        
```

The most substantiated complaints by far are found in the seventy-fifth precinct in East New York, Brooklyn, the location of a major corruption scandal, but nearby neighborhoods in Brooklyn also see a disproportionate number of complaints, as does the South Bronx. This map seems to show the results of over-policing minoritized communities.

The ProPublica database records the ethnicity of both the complainant and the accused officer. 

```{r}
cops <- allegations %>%
        group_by(unique_mos_id) %>%
        summarise(ethnicity = mos_ethnicity,
                  type = "Accused Officers") %>%
        distinct(.keep_all = TRUE)

ethnicity <- allegations %>%
        group_by(complaint_id) %>%
        summarise(ethnicity = complainant_ethnicity,
                  type = "Complainants") %>%
        distinct(.keep_all = TRUE) %>%
        rbind(cops) %>%
        filter(!(is.na(ethnicity)),
               ethnicity != "Unknown",
               ethnicity != "Refused")

remove(cops)

ggplot(ethnicity) +
        geom_bar(aes(ethnicity, fill = ethnicity)) +
        facet_wrap(~ type, scales = "free_x") +
        coord_flip() +
        labs(title = "Complaints Against NYPD by Ethnicity, 1985 to 2020",
             subtitle = "Most accused officers are white, most complainants are Black",
             caption = "Source: ProPublica",
             x = NULL,
             y = NULL) +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              axis.ticks.y = element_blank(),
              legend.position = "none") +
        scale_fill_brewer()
        
        
```

A majority of officers who have substantiated complaints against them are white while, in cases where the ethnicity is known, a majority of complainants are Black.

